---
title: "Methods and Results 14 December 2015"
author: "Daniel Wall"
date: "December 14, 2015"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}

setwd("/git_repositories/timeratios/wave_10dec/")
source("load_packages.R")
rm(list = ls())

load("analyzed_data/analyzed_data.Rdata")
# opts_chunk$set(echo=FALSE) 
```


# Summary

We are trying to see if diminishing impatience increases to a greater exetent when ratios change a lot. 


# Unlikeness scores

We want to see a bigger change in perceived differences between items when the ratio change is large as compared to small.

I subtracted the not delayed onset from the delayed onset. 

## Mean and CI of differences 

```{r, echo=FALSE}
ggplot(disssum, aes(x = ratcng, y = mean_diffcng)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper)) +
  ylab("Difference between not delayed and delayed intervals")
  
```

## Model predicting change in unlikeness

```{r, echo=FALSE}
Anova(diss_lmer, type = 3)
summary(diss_lmer)
```

### Plot of effects

```{r, echo=FALSE}
ggplot(eff_df, aes(x = ratcng, y = fit)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper))
```


# Intertemporal choices 

We want there to be a large ratio change in diminishing impatience for the large ratio change as compared to the small ratio change.

## Graph of mean percent larger later

```{r, echo=FALSE}

ggplot(itc_sum,
       aes(x = diss_set, y = estimate, group = delayed, fill = delayed, 
           ymax = conf.high,  ymin = conf.low)) +
  geom_bar(stat="identity", position = "dodge") +
  ylab("Percent Larger Later") +
  geom_errorbar(position = "dodge")
  
```


## Logistic regression predicting choices

```{r, echo=FALSE}
Anova(itc_lmer, type=3)
summary(itc_lmer)
```

### Effects plot from logistic regression 

```{r, echo=FALSE}
ggplot(itc_eff, aes(x = diss_set, y = fit, group = delayed,
                    fill = delayed,
                    ymin = lower, ymax = upper)) +
  geom_bar(stat="identity", position = "dodge") +
  geom_errorbar(position = "dodge") +
  ylab("Logistic Regression Coefficient")

```


