---
title: "ITCH Predictions"
author: "Daniel Wall"
date: "November 17, 2015"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---

From @MarzilliEricson2015

According to ITCH the probability of choosing larger later is


$$P(LL) = L \left(\beta_I + \beta_{xA}(x_2 - x_1) + \beta_{xR} \frac{x_2 - x_1}{x^*} + \beta_{tA}(t_2 - t_1) + \beta_{tR} \frac{t_2 - t_1}{t^*}\right)$$


Where $x^* = (x_1 + x_2)/2$


The function below calculates the probability that a participant will choose larger later.  

```{r}
p_ll <- function(beta_intercept, beta_xa, beta_xr, beta_ta, beta_tr, x1, x2, t1, t2) {
  x_rat <- (x2 - x1)/ ((x2 + x1)/2)
  t_rat <- (t2 - t1)/ ((t2 + t1)/2)
  reg <- beta_intercept + beta_xa*(x2 - x1) + beta_xr*(x_rat) + 
    beta_ta*(t2 - t1) + beta_tr*(t_rat)
  plogis(reg)
} 
```


To confirm that it is right let's use the example from the paper

```
Consider the following concrete example. If the par- ticipant uses weights (0.0, 0.1, 0.1, −0.1, −0.1) and selects between $10 today and $20 tomorrow, then the probabil- ity of choosing the later option is .68
```

```{r}
p_ll(0.0, 0.1, 0.1, −0.1, −0.1, 10, 20, 0, 1)
```



The function appears to be working



Let's see what is says for our example of $10 in 5 days and $11 in 6 days




```{r}
p_ll(0.0, 0.1, 0.1, −0.1, −0.1, 10, 11, 5, 6)

```


Now if we add 5 days to both 


```{r}

p_ll(0.0, 0.1, 0.1, −0.1, −0.1, 10, 11, 10, 11)

```


$P_{LL}^{10, 11}$   from the $P_{LL}^{5, 6}$ `r p_ll(0.0, 0.1, 0.1, −0.1, −0.1, 10, 11, 10, 11) - p_ll(0.0, 0.1, 0.1, −0.1, −0.1, 10, 11, 5, 6)`.



Now if we use $10 in 5 days and $11 in 15 days'


```{r}
p_ll(0.0, 0.1, 0.1, −0.1, −0.1, 10, 11, 5, 15)

```

Now by adding 5 days to each

```{r}
p_ll(0.0, 0.1, 0.1, −0.1, −0.1, 10, 11, 10, 20)

```

When subtracting the $P_{LL}^{10, 20}$   from the $P_{LL}^{5, 15}$ we get `r p_ll(0.0, 0.1, 0.1, −0.1, −0.1, 10, 11, 10, 20) - p_ll(0.0, 0.1, 0.1, −0.1, −0.1, 10, 11, 5, 15)`

So ITCH would predict that the there would be a small  difference in decreasing impatience between the two scenarios. I think we weren't accounting for the difference earlier today. 