---
title: "Scholten Read Sanborn"
author: "Daniel Wall"
date: "November 12, 2015"
output: 
  html_document: 
    number_sections: yes
    toc: yes
runtime: shiny
---

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(shinyapps)
library(dplyr)
```


```{r, echo=FALSE}
prob_dist <- function (p, alpha) {
  exp(-(-log(p))^alpha)
}

prob_df <- data.frame(prob = seq(0, 1, .01))

```

# Exponential Discounting

In the exponential discounting model, the decision maker will be indifferent between SS and LL when

$$\delta^{t_S}x_S = \delta^{t_L}x_L$$


Where $0 < \delta < 1$ is a one period discount factor and $\rho = 1 - \delta$

One period discount factor, an empiricial measure of discounting over that interval:

$$\delta = (\frac{x_S}{x_L})^{\frac{1}{t_L - t_S}}$$


how does the on period discount factor change for different values  of smaller sooner amount/time, and larger later amount/time. 

```{r, echo=FALSE}
exp_disfac <- function(ssamt, sstime, llamt, lltime) {
  (ssamt/llamt)^(1/(lltime - sstime))
}
```



```{r, echo=FALSE}
inputPanel(
  sliderInput("ssamt", "Smaller Sooner $", min=0, max = 500, step=5, value= 10),
  sliderInput("llamt", "Larger Later $", min=0, max = 500, step=5, value= 20),
  sliderInput("sstime", "Smaller Sooner time", min=0, max = 365, step=1, value= 1),
  sliderInput("lltime", "Larger Later time", min=0, max = 365, step=1, value= 2)
  )

renderPrint({
exp_disfac(input$ssamt, input$sstime, input$llamt, input$lltime) 
})
```

# Hyperbolic


According to Loewenstein and Prelec’s (1992) hyperbolic discounting model, the deci-
sion maker will be indifferent between SS and LL when 

$$d(t_S)v(x_S) = d(t_L)v(x_L)$$

where v is a reference-dependent value function, and d is a generalized hyperbolic dis- count function:

$$d(t) = (1 + \tau t)^{-\beta/\tau}$$

where $\beta > 0$ is delay discounting, and $\tau > 0$ is the departure from exponential discounting. As $\tau$ goes to 0, the generalized hyperbolic discount function reduces to the exponential discount function $d(t)\exp^{-\beta t} = \delta^t$. The hyperbolic discounting model predicts the common difference effect. To see this, the total discounting over the interval tS ? tL can be expressed in the form of a discount **fraction**. 

__So discount fraction is different than discount factor__


$$\frac{v(x_S)}{v(x_L)} = (\frac{1+\tau t_S}{1+\tau t_L}) ^{\beta/\tau} = (\frac{1+\tau t_S}{1 + \tau t_S + \tau(t_L-t_x)}) ^{\beta /\tau}$$


```{r}
hyp_disfrac <- function(sstime, lltime, beta, tau) {
  sstime <- as.numeric(sstime)
  lltime <- as.numeric(lltime)
  beta <- as.numeric(beta)
  tau <- as.numeric(tau)
  
  ((1 + tau * sstime)/(1 + tau * sstime + tau *(lltime - sstime))) ^(beta/tau)
}

hyp_disfrac(0, 1, .1, .1)
```


```{r, echo=FALSE}
inputPanel(
  textInput("sstime", "Smaller Sooner time"),
  textInput("lltime", "Larger Later time"),
  textInput("beta", "beta > 0 delay"),
  textInput("tau", "tau > departure from exponential discounting")
  )

renderPrint({
hyp_disfrac(input$sstime, input$lltime, input$beta, input$tau)

  
})

```


The value function has 5 properties. First three from Prospect theory

1. Reference Dependence
2. Diminishing Sensitivity
3. Loss aversion

2 additional dealing with elasticity of value function or the ratio between the proportional change in value and the proportional change in amount

$$\epsilon = \frac{v' (x)/v(x)}{1/x}$$

4. Increasing elasticity. For constant proportional increases of x > 0 (decreases of x < 0), v(x) increases (decreases) by increasing proportional amounts. Therefore, if outcome magnitude is increased by a common multiplicative constant, the ratio between the values of xS and xL decreases

5. Loss amplification. Elasticity is greater for losses than for gains. That is, decreasing x < 0 by a given proportion yields a greater propor- tional change in value than increasing x > 0 by the same proportion


```{r}
hyp_val <- function(x, gamma, lambda, mu, sigma) {
  # x <- as.numeric(x)
  if (x >= 0) {
    (1 - gamma) * x^(1-gamma) + (mu *gamma * x) ^ gamma
  } else if (x < 0) {
    -lambda*((1-gamma)*(-x) ^(1-gamma) + (mu + sigma)*gamma*(-x)^gamma)
  } 
}

hyp_val <- function(x, gamma, lambda, mu, sigma) {
  # x <- as.numeric(x)
  ifelse(x >= 0,(1 - gamma) * x^(1-gamma) + (mu *gamma * x) ^ gamma, -lambda*((1-gamma)*(-x) ^(1-gamma) + (mu + sigma)*gamma*(-x)^gamma))
  
}

hyp_val_df <- data.frame(x = -1000:1000)

hyp_val_df %>% rowwise %>% mutate(value = hyp_val(x, 1, 1, 1, 1))
hyp_val(-10, 1, 1, 1, 1)

```

where ½< c < 1 is diminishing absolute sensitivity to outcomes, l > 0 is increasing elasticity, k > 1 is loss aversion, and r > 0 is loss amplification



```{r, echo=FALSE}
inputPanel(
  sliderInput("gamma", "gamma diminishing absolute sensitivity to outcomes",  min=.1, max = .99, step=.001, value= .1),
  sliderInput("lambda", "Lambda loss aversion",  min=1, max = 4, step=.001, value= .1),
  sliderInput("mu", "mu  increasing elasticity",  min=.1, max = .99, step=.001, value= .1),
  sliderInput("sigma", "sigma loss amplification",  min=.1, max = .99, step=.001, value= .1)
  )


renderPlot({
  hyp_val_df$lambda <- input$lambda
  hyp_val_df$sigma <- input$sigma
  hyp_val_df$mu <- input$mu
  hyp_val_df$gamma <- input$gamma
  
  hyp_val_df <- transform(hyp_val_df, value = hyp_val(x, gamma, lambda, mu, sigma))
  
  ggplot(hyp_val_df, aes(x = x, y = value, group = 1)) +
    geom_line() +
    geom_abline(intercept = 0, slope=0) +
    geom_vline(xintercept = 0) +
    xlab("Outcome") +
    ylab("Value") +
    xlim(c(-1000, 1000))
})



```


# The interval discounting model: A hybrid evaluation rule 

uses the value function hyperbolic

Interval discount function:

$$D(t_S, t_L) = (1 + \alpha (\frac{w(t_L) - w(t_S)}{\vartheta})^{\vartheta})^{-\beta/\alpha}$$


```{r, echo=FALSE}
w_t <- function (t, tau) {
  t <- as.numeric(t)
  tau <- as.numeric(tau)
  (1/tau) * log(1 + tau*t)
}

int_fn <- function(int_len, int_onset, alpha, theta, beta, tau) {

  int_len <- as.numeric(int_len)
  int_onset <- as.numeric(int_onset)
  alpha <- as.numeric(alpha)
  theta <- as.numeric(theta)
  beta <- as.numeric(beta)
  
  sstime <- int_onset
  lltime <- sstime + int_len
  
  (1 + alpha * ((w_t(lltime, tau) - w_t(sstime, tau))/theta)^theta) ^(-beta/alpha)
}


int_df <- data.frame(int_len = 1:100)

int_df1 <- int_df %>%
  mutate(int_onset = 1, 
         alpha = .1, theta = .1, beta = .1, tau = .1)
# 
# mutate(int_df1, delta = int_fn(int_len, int_onset, alpha, theta, beta, tau))

```


```{r, echo=FALSE}
inputPanel(
  sliderInput("inton", "Interval onset", min=0, max = 365, step=1, value= 10),
  sliderInput("alpha", "alpha > 0 subadditivity", min=.01, max = 2, step=.05, value= 0),
  sliderInput("beta", "beta > 0 delay", min=0, max = 1, step=.05, value= .5),
  sliderInput("theta", "theta > 1 superadditive", min=1, max = 3, step=.05, value= 1),
  sliderInput("tau", "tau > 0 diminishing absolute sensitivity", min=.001, max = 1, step=.05, value= .01)
  )

renderPlot({
  int_df$inton <- input$inton
  int_df$alpha <- input$alpha
  int_df$beta <- input$beta
  int_df$theta <- input$theta
  int_df$tau <- input$tau
  
  int_df <- transform(int_df, delta = int_fn(int_len, inton, alpha, theta, beta, tau))
  
    int_df <- transform(int_df, daily = delta ^ (1/int_len))
  
  ggplot(int_df, aes(x = int_len, y = daily, group = 1)) +
    geom_line() +
#     geom_abline(intercept = 0, slope=0) +
#     geom_vline(xintercept = 0) +
#     xlab("Outcome") +
#     ylab("Value") +
    ylim(c(0, 1))
})

```


# Tradeoff model

attribute-based evaluation rule, in which the outcome advantage of one
option is weighted against the time advantage of the other, and the option with the great- est advantage is chosen.

These advantages are differences between weighted delays, w(tL) ? w(tS), or the effective interval (as in the interval discounting model), and differences between valued outcomes,
$v(x_L) - v(x_S)$ for $x_L > x_S > 0$ and $v(x_S) - v(x_L)$ for $x_L < x_S < 0$, or the effective compensation. The effective interval and the effective compensation are weighted against one another by a tradeoff function Q. 


                
$$\begin{equation}
 Q(w(t_L) - w(t_S)) = 
    \begin{cases}
      v(x_L) - v(x_S), & \text{if}\ x_L > x_S > 0 \\
      v(x_S) - v(x_L), & \text{if}\ 0 > x_S > x_L
    \end{cases}
  \end{equation}$$

Analogous to the time-weighing function, the trade- off model has a normalized logarithmic value function:

$$\begin{equation}
 v(x) = 
    \begin{cases}
      \frac{1}{\gamma}\log(1+\gamma x), & \text{if}\ x > 0 \\
      -\frac{\lambda}{\gamma}  \log(1+\gamma (-x)), & \text{if}\ x < 0
    \end{cases}
  \end{equation}$$



where $\gamma > 0$ is diminishing absolute sensitivity to outcomes. Like the normalized expo- nential value function discussed by K
obberling and Wakker (2005), the value function in
Eq. 10 approaches constant sensitivity (unit elasticity) as $\gamma$ goes to 0 (linear value func- tion), approaches insensitivity (zero elasticity) as $\gamma$ goes to infinity (zero value function), and is decreasingly elastic otherwise, spanning the entire range from unit elasticity (as x goes to 0) to zero elasticity (as x goes to infinity)


```{r,echo=FALSE}
to_val_fn <- function(x, gamma, lambda) {
  x <- as.numeric(x)
  gamma <- as.numeric(gamma)
  lambda <- as.numeric(lambda)
  
  ifelse(x >= 0, 
         (1/gamma) * log(1 + gamma * x),
         -(lambda/gamma) * log(1+gamma*(-x)))
}

```



```{r,echo=FALSE}
inputPanel(
  sliderInput("gamma_to", "gamma diminishing absolute sensitivity to outcomes",  min=.01, max = .99, step=.001, value= .1),
  sliderInput("lambda_to", "Lambda loss aversion", value = 1,  min=1, max = 4, step = .05)
)


renderPlot({
  to_val_df <- data.frame(x = -100:100)
  to_val_df$lambda <- input$lambda_to
  to_val_df$gamma <- input$gamma_to
  


  to_val_df <- mutate(to_val_df, value = to_val_fn(x, gamma, lambda))
  
  ggplot(to_val_df, aes(x = x, y = value, group = 1)) +
    geom_line() +
    geom_abline(intercept = 0, slope=0) +
    geom_vline(xintercept = 0) +
    xlab("Outcome") +
    ylab("Value") +
    xlim(c(-100, 100))
})
```


The tradeoff fucntion weighs effective intervals against effective compensations

$$Q(t_S, t_L) = \frac{\kappa}{\alpha} \log
\left(1 + \alpha \left(\frac{w(t_L) - w(t_S)}{\vartheta}\right)^{\vartheta}\right)$$
